Research paper

### 2013-01-24 Research paper.tex
RP: Added suggestion by Maarten

\subsection{The builder: machine translation} \label{builder}
Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology. A project overview therefore cannot be complete without an overview of the software's functionality. Colibri's functionality can be split in four parts: (1) identifying and extracting constructions in a language, (2) identifying aligned constructions between language-pairs, (3) memory-based classification using construction experts and (4) reassembling constructions into a coherent target sentence. I will discuss the four parts in more detail consecutively:

@@

\subsubsection{Memory-based classification using construction experts}
As may have become clear from the previous sections, the Colibri project investigates the use of machine learning techniques to get from the source language construction to the target language constructions. More specifically, Colibri uses specialized classifiers, trained for only one construction, inspired by the effectivity of this approach in word sense ambiguation \citep{vangompel10}. The results of this approach are compared to the results of a combined classifier approach and statistical machine translation.

@@

At the moment, the functional machine translation system described in section \ref{builder} is mostly finished. It is mainly inspired upon the open source translation system MOSES, and indeed manages to produce similar results. However, as was explained in section \ref{Description}, Colibri is an attempt to improve the quality of machine translation by adding skipgrams and memory-based learning. Unfortunately it has turned out that both do not add much to the translation quality: memory-based learning only improves the results a little bit, and the skipgrams are mostly ignored by the decoder. Apparently, translation hypotheses without skipgrams are more likely to be good translations in most of the cases. \\\indent
This means Colibri's goal, improving translation quality, has not yet been achieved. It is possible that larger improvements in translation quality \emph{can} be achieved by changing things like corpora or even languages used, but if this is not the case, new possibilities for improvements have to be investigated. One such possibility is to include a system that finds the correct translations of words on the basis of particular keywords in a sentences. This system should for example be able to translate the English \emph{bank} into Dutch \emph{oever} ('river bank') when the words like \emph{boat} are mentioned, but into \emph{bank} (financial institution) when \emph{money} is mentioned.

>>>>

\subsection{The builder: machine translation} \label{builder}
Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology. A project overview therefore cannot be complete without an overview of the software's functionality. Colibri's functionality can be split in four parts: (1) identifying and extracting constructions in a language, (2) identifying aligned constructions between language-pairs, (3) memory-based classification and (4) reassembling constructions into a coherent target sentence. I will discuss the four parts in more detail consecutively:

@@

\subsubsection{Memory-based classification}
As may have become clear from the previous sections, the Colibri project investigates the use of machine learning techniques to get from the source language construction to the target language constructions. Three approaches have been implemented:

\begin{itemize}
\item Monolithic classifier. All phrases go through one general classifier. This classifier thus has a very large number of classes to pick from.
\item N-array classifiers. There is a separate classifier for each phrase length (n); the length of the source language construction thus decides which classifier is used.
\item Construction experts. Each construction or phrase has its own classifier, which means that instead of one larger classifier, there are a lot of smaller classifier making a simpler decision. This is inspired by the effectivity of this approach in word sense ambiguation \citep{vangompel10}.
\end{itemize}

@@

At the moment, the functional machine translation system described in section \ref{builder} is mostly finished. It is mainly inspired upon the open source translation system MOSES, and indeed manages to produce similar results. However, as was explained in section \ref{Description}, Colibri is an attempt to improve the quality of machine translation by adding skipgrams and memory-based learning. Unfortunately it has turned out that both do not add much to the translation quality: all three memory-based learning approaches (construction experts, N-array classifiers and a monolithic classifier) only improve the results a little bit, and the skipgrams are mostly ignored by the decoder. Apparently, translation hypotheses without skipgrams are more likely to be good translations in most of the cases. \\\indent
This means Colibri's goal, improving translation quality, has not yet been achieved. It is possible that larger improvements in translation quality \emph{can} be achieved by changing things like corpora or even languages used, but if this is not the case, new possibilities for improvements have to be investigated. One such possibility is to include a system that finds the correct translations of words on the basis of particular keywords in a sentences. This system should for example be able to translate the English \emph{bank} into Dutch \emph{oever} ('river bank') when the words like \emph{boat} are mentioned, but into \emph{bank} (financial institution) when \emph{money} is mentioned.
